### 2023-10-11 codebook.qmd
Merge remaining files from dataset repo

---
title: Codebook v0.1
author: J.L.A. Krusell
format:
    pdf:
        header-includes: \renewcommand{\arraystretch}{1.7}
---

Introduction, coverage

only includes frozen onsets, not all outcomes

only includes first onset

### Definitions

Define a dyadic conflict episode

### Variables

| Variable      | Type    | Description                                                 |
|:--------------|:--------|:------------------------------------------------------------|
| `conflict_id` | Integer | Unique conflict-level identifier                            |
| `dyad_id`     | Integer | Unique dyadic-level identifier                              |
| `year`        | Integer | Onset year                                                  |
| `duration`    | Integer | Number of years until the conflict becomes active again     |
| `side_a`      | String  | Primary state actor in the dyadic-conflict                  |
| `side_b`      | String  | Primary non-state actor in a conflict dyad                  |
| `frozen`      | Boolean | Binary value indicating an onset of a frozen civil conflict |
| `comment`     | String  | Coding notes                                                |
| `confidence`  | Float   | Self-reported coding confidence falling within $[0.5, 1]$   |