### 2023-11-28 slides/gues3s-23-11-27.qmd
guess3s presentation

---
title: "Frozen Civil Conflicts"
subtitle: "The Role of External State Support"
author:
    - name: J.L.A. Krusell
institute: "Department of Political Science"
format:
    revealjs:
        theme: solarized
        footer: "Department of Political Science --- University of Gothenburg"
suppress-bibliography: true
---

```{r setup, include F}
library(dplyr)
library(fc.utils)
library(ggdist)
library(ggplot2)
library(ggthemes)
library(tidyr)

theme_set(theme_solarized())
```

## Defining Frozen Conflicts

"...protracted, post-war conflict process, characterized by the absence of stable peace...[where] core issues between the opposing sides remain unresolved, the dispute is in the forefront of mutual relations, and there is a looming threat of the renewal of violence." [@smetana_between_2019]

## Research Question {auto-animate=true}

Why do some civil conflicts become frozen?

## Research Question {auto-antimate=true}

To what extent does external state support affect the likelihood of a civil conflict becoming frozen?

. . .

::: {.absolute bottom=50%}
$\text{External State Support} \rightarrow \text{Frozen Civil Conflict Onset}$
:::

## Methods

Two strategies: bayesian probit model and instrumental variable analysis

. . .

- Unit of Analysis: Civil conflict episodes [@gleditsch_armed_2002]

- Outcome: Frozen conflict onset

- Treatment:
  1. External state support for *rebels*
  2. External state support for *government*
  3. Interaction

## Measuring Frozen Conflicts

Civil conflict episodes were coded as becoming frozen conflicts if the following conditions were satisfied:

- Episode did **not** terminate with either a one-sided military victory or comprehensive peace agreement

- Non-state actor continued to exist for at least 5 years following episode termination without renouncing political violence

## Measuring Frozen Conflicts

248 civil conflict episodes spanning 1975-2017, of which 69 became frozen

```{r}
dataset <- readRDS("../data/model_data.rds")

filter(dataset, frozen == 1) |>
ggplot(aes(year)) +
    geom_bar() +
    xlab("Year") +
    ylab("Frozen Civil Conflict Onsets")
```

## Measuring Frozen Conflicts {.smaller}

```{r}
conditions <- dataset |>
    mutate(treatment = case_when(ext_sup_s_rebel_bin == 1 & ext_sup_s_state_bin == 1 ~ "Ext. Sup. to Both",
                                 ext_sup_s_rebel_bin == 1 & ext_sup_s_state_bin == 0 ~ "Ext. Sup. Rebel Only",
                                 ext_sup_s_rebel_bin == 0 & ext_sup_s_state_bin == 1 ~ "Ext. Sup. Gov Only",
                                 ext_sup_s_rebel_bin == 0 & ext_sup_s_state_bin == 0 ~ "Ext. Sup. to Neither"))

m <- table(conditions$frozen, conditions$treatment)
rownames(m) <- c("Not Frozen", "Frozen")
knitr::kable(m)
```

## Additional Covariates

:::{.incremental}
- Conflict dynamics: duration, intensity, recurrence
- Incompatibility
- Ongoing intra/inter-state conflict
- Cold war
- Censoring
:::

## Probit Model

\begin{align*}
y^\star_i & = \alpha + X_i' \beta + T_i' \delta_{k_i} + \gamma_{c_i} + \epsilon_i, \: \epsilon_i \sim \text{N}(0, 1) \\
y_i & = \begin{cases}
1 & \text{if } y_i^\star > 0 \\
0
\end{cases}
\end{align*}

. . .

:::: {.columns}

::: {.column width="30%"}
\begin{align*}
\alpha & \sim \text{N}(0, 2.5) \\
\beta & \sim \text{N}(0, 1) \\
\end{align*}
:::

::: {.column width="40%"}
\begin{align*}
\delta_j & \sim \text{N}(\mu_j, \tau_j) \\
\mu_j & \sim \text{N}(0, 1) \\
\tau_j & \sim \text{N}_{+}(0, 1)
\end{align*}
:::

::: {.column width="30%"}
\begin{align*}
\gamma_c & \sim \text{N}(0, \sigma) \\
\sigma & \sim \text{N}_{+}(0, 1)
\end{align*}
:::

:::

## Parameter Estimates

```{r}

fit <- readRDS("../posteriors/frozen-bin-all/fit.rds")

delta <- fit$draws("delta", format = "df") |>
    select(matches("delta")) |>
    pivot_longer(everything(), names_to = "variable")

ll <- extract_digits(delta$variable)
delta$variable <- sprintf("delta[%s]", sapply(ll, `[`, 1))
delta$group <- ifelse(sapply(ll, `[`, 2) == 1, "Government Incompatibility",
                      "Territorial Incompatibility")

ggplot(delta, aes(x = value, y = variable, fill = variable)) +
    stat_halfeye(alpha = 0.8, show.legend = F) +
    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.2) +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank()) +
    scale_y_discrete(labels = c("delta[1]" = "Ext. Sup. to Gov",
                                "delta[2]" = "Ext. Sup. to Rebel",
                                "delta[3]" = "Interaction")) +
    facet_grid(. ~ group) +
    scale_fill_solarized()

```

## Average Marginal Effects {.smaller}

Three treatment conditions:

#. External state support for the **government only**
#. External state support for the **rebels only**
#. External state support for **both sides**

\begin{equation}
\text{AME}_{jk} = \frac{1}{N_k} \sum_{i=1}^{N_k} \Phi(\alpha + X_i' \beta + W_j' \delta_k + \gamma_{c_i}) - \Phi\left(\alpha + X_i' \beta + \gamma_{c_i}\right)
\end{equation}

## Average Marginal Effects

```{r}
ame <- fit$draws("ame", format = "df") |>
    select(matches("ame")) |>
    pivot_longer(everything(), names_to = "variable")

ll <- extract_digits(ame$variable)
ame$variable <- sprintf("ame[%s]", sapply(ll, `[`, 1))
ame$group <- ifelse(sapply(ll, `[`, 2) == 1, "Government Incompatibility",
                    "Territorial Incompatibility")

ggplot(ame, aes(x = value, y = variable, fill = variable)) +
    stat_eye(alpha = 0.8, show.legend = F) +
    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.2) +
    theme(axis.title.x = element_blank()) +
    labs(x = parse(text = "bold('AME '*Delta)")) +
    scale_y_discrete(labels = c("ame[1]" = "Ext. Sup. Gov Only",
                                "ame[2]" = "Ext. Sup. Rebel Only",
                                "ame[3]" = "Ext. Sup. to Both")) +
    coord_flip() +
    facet_grid(. ~ group) +
    scale_fill_solarized()

```

# Part II

## Instrument

UN General Assembly voting 10 years prior to a conflict onset

. . .

**Pros**:

- No measurement error, unlikely to have a direct effect on civil conflict

. . .

**Cons**:

- Cannot exclude possibility of additional systematic confounders

## I.V. Model

\begin{align*}
T^\star_i & = \alpha_1 + X_i' \beta_1 + Z_i' \lambda + \upsilon_{c_i} + \eta_i \\
y^\star_i & = \alpha_2 + X_i' \beta_2 + T_i' \delta_{k_i} + \gamma_{c_i} + \epsilon_i \\
\eta_i, \epsilon_i & \sim \text{MVN}(0, \Sigma)
\end{align*}