### 2023-11-30 slides/peer-presentation-23-11-30.qmd
Polisci peer presentation

---
title: "Frozen Civil Conflicts"
subtitle: "The Role of External State Support"
author:
    - name: J.L.A. Krusell
institute: "Department of Political Science"
format:
    revealjs:
        theme: solarized
        footer: "Department of Political Science --- University of Gothenburg"
        title-slide-attributes:
            data-background-image: ../assets/LO_GUeng_CMYK_flag.png
            data-background-size: 10%
            data-background-position: 2% 0%
suppress-bibliography: true
---

```{r setup, include F}
library(dplyr)
library(fc.utils)
library(ggdist)
library(ggplot2)
library(ggthemes)
library(tidyr)

theme_set(theme_solarized())
```

## Introduction {.smaller}

*To what extent does external state support affect the likelihood of a civil conflict becoming frozen?*

::: {.incremental}

- Previous research on the link between external support and conflict termination [@mason1996civil; @balch2008third; @karlen2017legacy; @sawyer2017role]

    - Generally assumes two possible outcomes: military victory or peace agreement

    - Very little scholarly work on frozen conflicts

- **My contribution**:

    - Theoretical, extending bargaining theory to the domain of frozen conflicts
    - Empirical, novel dataset to examine relationship between external support and frozen civil conflict onset

:::

## Defining Frozen Conflicts

"...protracted, post-war conflict process, characterized by the absence of stable peace...[where] core issues between the opposing sides remain unresolved, the dispute is in the forefront of mutual relations, and there is a looming threat of the renewal of violence." [@smetana_between_2019]

# Theory

## Government Contestation

::: {.incremental}

- External support shifts the balance of power

    - Often assumed that the rebel side is the weaker party in civil conflicts

- Introduces additional veto players that complicates peace negotiations

:::

## Territorial Contestation {.smaller}

::: {.incremental}

**For the rebels**:

- Opportunity to consolidate power, and establish a *de facto* state free from credible committments issues arising from negotiated solutions

    - External support should be decisive in aiding rebels to seize territory

**For the government**:

- Future prospect of forcible reintegration

    - External support extends time horizon and offsets accumulated costs

:::

# Empirics

## Methods

Two strategies: bayesian probit model and instrumental variable analysis

. . .

- Unit of Analysis: Civil conflict episodes [@gleditsch_armed_2002]

- Binary outcome: Frozen conflict onset

- Binary treatments:
  1. External state support for *rebels*
  2. External state support for *government*
  3. Interaction

## Measuring Frozen Conflicts

Civil conflict episodes were coded as becoming frozen conflicts if the following conditions were satisfied:

- Episode did **not** terminate with either a one-sided military victory or comprehensive peace agreement

- Non-state actor continued to exist for at least 5 years following episode termination without renouncing political violence

## Measuring Frozen Conflicts {.smaller}

248 civil conflict episodes spanning 1975-2017, of which 69 became frozen

:::: {.columns}

::: {.column}
```{r}
#| fig-height: 7
dataset <- readRDS("../data/model_data.rds")
frozen <- filter(dataset, frozen == 1) |>
    mutate(country = sub("Government of ", "", side_a),
           country = case_when(country == "Russia (Soviet Union)" ~ "Russia",
                               country == "United Kingdom" ~ "UK",
                               country == "Congo" ~ "Republic of Congo",
                               country == "Myanmar (Burma)" ~ "Myanmar",
                               country == "Bosnia-Herzegovina" ~ "Bosnia and Herzegovina",
                               country == "Serbia (Yugoslavia)" ~ "Serbia",
                               country == "DR Congo (Zaire)" ~ "Democratic Republic of the Congo",
                               T ~ country))

sub.df <- group_by(frozen, country) |> count(frozen)

world <- map_data("world") |>
    left_join(sub.df, by = c("region" = "country")) |>
    mutate(n = ifelse(is.na(n), 0, 1) )

ggplot(world, aes(long, lat, map_id = region, fill= n, group = group)) +
    geom_map(map = world) +
    theme(legend.position = "none") +
    scale_color_solarized()
```
:::

::: {.column}
```{r}
#| fig-height: 7
filter(dataset, frozen == 1) |>
ggplot(aes(year)) +
    geom_bar() +
    xlab("Year") +
    ylab("Frozen Civil Conflict Onsets")
```
:::
::::

## Measuring Frozen Conflicts {.smaller}

```{r}
conditions <- dataset |>
    mutate(treatment = case_when(ext_sup_s_rebel_bin == 1 & ext_sup_s_state_bin == 1 ~ "Ext. Sup. to Both",
                                 ext_sup_s_rebel_bin == 1 & ext_sup_s_state_bin == 0 ~ "Ext. Sup. Rebel Only",
                                 ext_sup_s_rebel_bin == 0 & ext_sup_s_state_bin == 1 ~ "Ext. Sup. Gov Only",
                                 ext_sup_s_rebel_bin == 0 & ext_sup_s_state_bin == 0 ~ "Ext. Sup. to Neither"))

m <- table(conditions$frozen, conditions$treatment)
rownames(m) <- c("Not Frozen", "Frozen")
knitr::kable(m)
```

## Additional Covariates

:::{.incremental}
- Conflict dynamics: duration, intensity, recurrence
- Incompatibility
- Ongoing intra/inter-state conflict
- Cold war
- Censoring
:::

# Probit Analysis

## Probit Model {auto-animate=true}

\begin{align*}
y^\star_i & = \alpha + X_i' \beta + T_i' \delta_{k_i} + \gamma_{c_i} + \epsilon_i, \: \epsilon_i \sim \text{N}(0, 1) \\
y_i & = \begin{cases}
1 & \text{if } y_i^\star > 0 \\
0
\end{cases}
\end{align*}

## Probit Model {auto-animate=true}

\begin{align*}
y^\star_i & = \alpha + X_i' \beta + T_i' \delta_{k_i} + \gamma_{c_i} + \epsilon_i, \: \epsilon_i \sim \text{N}(0, 1) \\
y_i & = \begin{cases}
1 & \text{if } y_i^\star > 0 \\
0
\end{cases}
\end{align*}

\begin{align*}
\alpha & \sim \text{N}(0, 2.5) \\
\beta & \sim \text{N}(0, 1) \\
\end{align*}

## Probit Model {auto-animate=true}

\begin{align*}
y^\star_i & = \alpha + X_i' \beta + T_i' \delta_{k_i} + \gamma_{c_i} + \epsilon_i, \: \epsilon_i \sim \text{N}(0, 1) \\
y_i & = \begin{cases}
1 & \text{if } y_i^\star > 0 \\
0
\end{cases}
\end{align*}

:::: {.columns}

::: {.column}
\begin{align*}
\alpha & \sim \text{N}(0, 2.5) \\
\beta & \sim \text{N}(0, 1) \\
\end{align*}
:::

::: {.column}
\begin{align*}
\delta_j & \sim \text{N}(\mu_j, \tau_j) \\
\mu_j & \sim \text{N}(0, 1) \\
\tau_j & \sim \text{N}_{+}(0, 1)
\end{align*}
:::

::::

## Probit Model {auto-animate=true}

\begin{align*}
y^\star_i & = \alpha + X_i' \beta + T_i' \delta_{k_i} + \gamma_{c_i} + \epsilon_i, \: \epsilon_i \sim \text{N}(0, 1) \\
y_i & = \begin{cases}
1 & \text{if } y_i^\star > 0 \\
0
\end{cases}
\end{align*}

:::: {.columns}

::: {.column width="30%"}
\begin{align*}
\alpha & \sim \text{N}(0, 2.5) \\
\beta & \sim \text{N}(0, 1) \\
\end{align*}
:::

::: {.column width="40%"}
\begin{align*}
\delta_j & \sim \text{N}(\mu_j, \tau_j) \\
\mu_j & \sim \text{N}(0, 1) \\
\tau_j & \sim \text{N}_{+}(0, 1)
\end{align*}
:::

::: {.column width="30%"}
\begin{align*}
\gamma_c & \sim \text{N}(0, \sigma) \\
\sigma & \sim \text{N}_{+}(0, 1)
\end{align*}
:::

::::

## Parameter Estimates

```{r}

fit <- readRDS("../posteriors/frozen-bin-all/fit.rds")

delta <- fit$draws("delta", format = "df") |>
    select(matches("delta")) |>
    pivot_longer(everything(), names_to = "variable")

ll <- extract_digits(delta$variable)
delta$variable <- sprintf("delta[%s]", sapply(ll, `[`, 1))
delta$group <- ifelse(sapply(ll, `[`, 2) == 1, "Government Incompatibility",
                      "Territorial Incompatibility")

ggplot(delta, aes(x = value, y = variable, fill = variable)) +
    stat_halfeye(alpha = 0.8, show.legend = F) +
    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.2) +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank()) +
    scale_y_discrete(labels = c("delta[1]" = "Ext. Sup. to Gov",
                                "delta[2]" = "Ext. Sup. to Rebel",
                                "delta[3]" = "Interaction")) +
    facet_grid(. ~ group) +
    scale_fill_solarized()

```

## Average Marginal Effects {.smaller}

Three treatment conditions:

#. External state support for the **government only**
#. External state support for the **rebels only**
#. External state support for **both sides**

\begin{equation}
\text{AME}_{jk} = \frac{1}{N_k} \sum_{i=1}^{N_k} \Phi(\alpha + X_i' \beta + W_j' \delta_k + \gamma_{c_i}) - \Phi\left(\alpha + X_i' \beta + \gamma_{c_i}\right)
\end{equation}

where $W_j$ is the corresponding treatment vector for treatment condition $j$.

## Average Marginal Effects

```{r}
ame <- fit$draws("ame", format = "df") |>
    select(matches("ame")) |>
    pivot_longer(everything(), names_to = "variable")

ll <- extract_digits(ame$variable)
ame$variable <- sprintf("ame[%s]", sapply(ll, `[`, 1))
ame$group <- ifelse(sapply(ll, `[`, 2) == 1, "Government Incompatibility",
                    "Territorial Incompatibility")

ggplot(ame, aes(x = value, y = variable, fill = variable)) +
    stat_eye(alpha = 0.8, show.legend = F) +
    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.2) +
    theme(axis.title.x = element_blank()) +
    labs(x = parse(text = "bold('AME '*Delta)")) +
    scale_y_discrete(labels = c("ame[1]" = "Ext. Sup. Gov Only",
                                "ame[2]" = "Ext. Sup. Rebel Only",
                                "ame[3]" = "Ext. Sup. to Both")) +
    coord_flip() +
    facet_grid(. ~ group) +
    scale_fill_solarized()

```

# I.V. Analysis

## Instrument {.smaller .nostretch}

:::: {.columns}

::: {.column}
UN General Assembly voting 10 years prior to a conflict onset

- Captures foreign policy homophily/heterophily
:::

::: {.column}
![](../assets/diagram.png)
:::

::::

## Instrument {.smaller auto-animate=true}

- Multiple approaches to measurement:

    - Average S-score (Signorino and Ritter 1999):

       \begin{align*}
        S_{ab} = 1 - \cfrac{\sum | Y_{av} - Y_{bv} |}{V}
        \end{align*}

        where $Y_{av}$ denotes the recorded vote for country $a$ on resolution $v$ for $v = 1, \ldots V$ within the target time period. Then, for a subset of countries, $\mathcal{F}$, relevant to country $a$,

        \begin{align*}
        Z_a = \frac{1}{|\mathcal{F}|} \sum_{i \in \mathcal{F}} S_{ai}
        \end{align*}

## Instrument {.smaller auto-animate=true}

- Multiple approaches to measurement:

    - IRT latent scores (Bailey et al 2017):

        \begin{align*}
        P(Y_{itv} = k) = \Phi(\gamma_{kv} - \beta_v \theta_{it}) - \Phi(\gamma_{k-1,v} - \beta_v \theta_{it})
        \end{align*}

        where $i = 1, \ldots, N$ now indexes the country. Then,

        \begin{align*}
        Z_a = \frac{1}{|\mathcal{F}| + |\mathcal{G}|} \sum_{i \in \mathcal{F}, t \in \mathcal{G}} \lVert \theta_{at} - \theta_{it} \rVert
        \end{align*}

        for the restricted time period $\mathcal{G}$.

## Instrument {.smaller}

:::: {.columns}

::: {.column width="60%"}

UN General Assembly voting

**Pros**:

- Unlikely to have a direct effect on civil conflict

:::

::: {.column width="40%"}
```{r}
#| fig-height: 4
#| fig-width: 4
library(dagitty)
library(ggdag)

dag <- dagitty("dag {
Z [pos=\"0,0.5\"];
T [pos=\"0,0\"];
Y [pos=\"1,0\"];
U [pos=\"0.5,-0.5\"];
Z -> T; T -> Y; U -> Y; U -> T
}")

ggdag(dag) +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_blank(),
          panel.background = element_blank())
```
:::

::::

## Instrument {.smaller}

:::: {.columns}

::: {.column width="60%"}
UN General Assembly voting

**Pros**:

- Unlikely to have a direct effect on civil conflict

**Cons**:

- Cannot exclude possibility of additional systematic confounders
- Difficult to instrumentize *both* treatments
:::

::: {.column width="40%"}
```{r}
#| fig-height: 4
#| fig-width: 4
library(dagitty)
library(ggdag)

dag <- dagitty("dag {
W [pos=\"0.5,1\"];
Z [pos=\"0,0.5\"];
T [pos=\"0,0\"];
Y [pos=\"1,0\"];
U [pos=\"0.5,-0.5\"];
Z -> T; T -> Y; U -> Y; U -> T; W -> Z; W -> Y
}")

ggdag(dag) +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text = element_blank(),
          axis.title = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_blank(),
          panel.background = element_blank())
```
:::

::::

## I.V. Model

\begin{align*}
T^\star_i & = \alpha_1 + X_i' \beta_1 + Z_i' \lambda + \gamma_{1c_i} + \eta_i \\
y^\star_i & = \alpha_2 + X_i' \beta_2 + T_i \delta + \gamma_{2c_i} + \epsilon_i \\
\eta_i, \epsilon_i & \sim \text{MVN}(0, \Sigma)
\end{align*}

Assumptions:

1. $\lambda \neq 0$
2. $Z \perp\!\!\!\perp Y \mid T$
3. $Z \perp\!\!\!\perp \epsilon \mid T, X$

# Conclusion

. . .

Should the instrument be completely abandoned?