### 2024-03-08 paper.qmd
First draft

---
title: "Frozen Civil Conflicts"
subtitle: "The Role of External State Support"
authors:
- name: J.L.A. Krusell
  email: "joshua.krusell@gu.se"
  affiliations:
      - "University of Gothenburg"
date: today
date-format: long
linestretch: 1.25
bibliography: library.bib
filters:
    - lib/luafilter/meta.lua
format:
    html:
        toc: true
    pdf:
        template-partials:
            - templates/before-body.tex
            - templates/title.tex
execute:
    cache: true
    echo: false
    warning: false
abstract: |
    Frozen conflicts are among some of the most pernicious, difficult to
    resolve conflicts. Surprisingly, little scholarly attention has been
    paid to the factors influencing their emergence. This paper addresses
    one possible dimension, namely the effect of external state support to
    conflict actors on the probability of an ongoing civil conflict
    becoming frozen. I use an original dataset coding frozen conflicts as
    an outcome among UCDP civil conflict episodes from 1975 to 2019 in
    order to construct a series of bayesian, hierarchical probit models
    motivated by an application of bargaining theory to the domain of
    frozen conflicts. Results indicate that directed support to rebels increases
    the probability that a civil conflict episode terminates as a frozen
    conflict within the context of territorial civil conflicts. All other
    treatment conditions --- specifically, directed
    support provided only to the government and directed supported given
    to both sides --- are found to be inconsequential. Taken together,
    these findings help provide a more nuanced understanding of conflict
    termination beyond the common assumption that conflicts either
    conclude through a negotiated settlement or a one-sided military
    victory. Furthermore, this paper also helps shed light on the
    conditions under which actors may prefer a sub-optimal solution,
    allowing a conflict to transition into a state of unstable peace with
    an ever present threat of violent re-escalation, rather than a
    conclusive termination.
---

[//]: # Setup R environment
```{r setup, include = F}
library(dplyr)
library(ggplot2)
library(ggthemes)
library(ggdist)
library(latex2exp)
library(knitr)
library(patchwork)
library(tidyr)

# Project specific utility package
library(fc.utils)

options(knitr.kable.NA = ".", knitr.kable.format = "latex",
        knitr.kable.linesep = "")

theme_set(theme_tufte(base_family = "TeX Gyre Pagella", base_size = 10))

design <- readRDS("./data/model_inputs/design.rds")
```

Frozen conflicts are among some of the most pernicious, difficult to resolve conflicts. They are unique among conflict outcomes in that they emerge from an inconclusive conflict termination, and leave conflict actors in an indeterminate state of unstable peace with an ever present threat of violent re-escalation.

Surprisingly, despite the geographic and temporal dispersion of frozen conflicts, little scholarly attention has been paid to the factors influencing their emergence [@klosek_frozen_2021]. The frozen conflict specific literature mostly consists of qualitative case studies of specific conflicts, while the broader academic work on conflict termination has so far not incorporated frozen conflict as a unique conflict outcome.

This paper seeks to address this gap by asking, why do some civil conflicts become frozen? There are a wide variety of potential factors which could plausibly be correlated with the emergence of frozen conflicts; however, I decide to focus on the international dimension, specifically the role of external support provided by third party state actors to conflict belligerents. There exists extensive work on the impact of external support on conflict onset, duration, and termination [@hartzell_civil_2016]. Although there may be disagreements on the exact relationship between external support and various aspects of civil conflicts, the aggregate finding is that external support is an important factor.

In what follows, I begin by providing an overview of the previous literature, both in terms of the frozen conflict specific work as well as the broader literature on civil conflict termination. I then offer a theory for why certain civil conflicts become frozen and link it to the role played by external support. In brief, conflicts freeze when actors seek to minimize the costs of fighting by de-escalating the conflict intensity when unable to either achieve a military advantage during fighting or agree to a negotiated settlement. External support is thought to be positively correlated with the likelihood that a civil conflict becomes frozen, primarily by shifting the balance of power under certain conditions towards a stalemate situation.

To empirically investigate this theory I provide an exploratory analysis using a series of bayesian hierarchical probit models with original coded data on frozen civil conflict onsets. The findings show that external state support does increase the probability that a civil conflict terminates in a frozen state, but only in the case of directed support to rebel groups in civil conflicts where the central incompatibility centers around territory. These results contribute to our understanding of frozen conflicts, as well as provide a more nuanced understanding of conflict termination more broadly.

# Literature Review

Previous empirical research into frozen conflicts is fairly limited. The work that does exist largely consists of qualitative, single-case studies [*e.g.*, @blank_russia_2008; @welt_thawing_2010; @pokalova_conflict_2015; @tudoroiu_unfreezing_2016]. While this literature provides richly detailed and illuminating analyses of specific frozen conflicts, it does not easily provide a comparative perspective to investigate commonalities and differences across cases. Thus, we know relatively little about the general correlates of frozen civil conflict emergence.

In contrast to the paucity of literature concerning frozen conflicts, there exists numerous studies examining the links between third party involvement on both the duration and outcome of civil conflicts [@shelton_what_2013]. Previous work has argued that external interventions by third party states increase the duration of civil conflicts [@regan_third-party_2002]. However, biased military interventions may be the exception and in some cases actually increase the probability of a favourable outcome for the supported conflict actor [@balch-lindsay_third-party_2008]. Although others have put forth empirical arguments that this only holds in the case of rebel biased interventions. More specifically, third party military interventions may decrease the duration of a conflict and increase the probability of either negotiated settlement or rebel military victory in part because interveners may only involve themselves in ongoing civil conflicts in favor of the government when the government is facing a relatively strong rebel group with a high level of military capabilities [@gent_going_2008]. The probability for a government military victory would on average be lower in these situations, thereby masking the effects of a government biased intervention. Indeed, when accounting for the differential effect of intervention given the strength of the opposition, there have been empirical claims that government biased interventions decrease the probability of a rebel victory and increase the probability of a government military victory [@sullivan_conditional_2015].

Additional work has sought to look at the relationships and preferences of supporting actors as well as the disaggregated types of external support. Third party states intervening or providing material support to opposing sides in a civil conflict for the purpose of balancing may increase conflict duration [@aydin_networks_2012; @anderson_competitive_2019]. When looking at disaggregated types of external interventions, some have argued that military aid and other forms of support are the key to either a rebel victory or negotiated settlement, as opposed to direct military involvement [@jones_altering_2017]. Others argue that fungible support, *i.e.* financing and weapons, has an opposite impact by actually extending the duration of conflicts [@sawyer_role_2017].

Taken together, this literature points to the importance of external support as a factor on both conflict duration and termination outcomes. Most studies tend to find positive support for the supposition that external support primarily benefits the rebel side, both in terms of increasing the chances of a military victory or a negotiated settlement. However, this may not hold when examining more fine grained measures for specific types of support offered to belligerents. Nonetheless, the common thread across these works is a general assumption that conflicts terminate either with a one-sided military victory or a negotiated settlement. Frozen conflicts, however, represent a qualitatively different type of conflict termination. Unlike a one-sided military victory, there is no clear victor whena conflict freezes, rather the conflict stales with both sides either unwilling or unable to continue large-scale hostilities, And, compared to a negotiated settlement, they crucially do not encompass comprehensive peace agreements which would otherwise conclusively terminate a conflict. Thus frozen conflicts represent a unique outcome; one that has previously not been examined quantitatively, especially in regards to international dynamics such as external state support.

# Theory

## Defining Frozen Conflict

I start with the definition of frozen conflict offered by @smetana_between_2019:

> We define frozen conflict as a protracted, post-war conflict process, characterized by the absence of stable peace... [where] core issues between the opposing sides remain unresolved, the dispute is in the forefront of mutual relations, and there is a looming threat of the renewal of violence.

Accompanying this definition, @smetana_between_2019 lay forth four criteria determining whether a post-conflict period qualifies as a frozen conflict. First, the conflict must be international since it is assumed that frozen conflict is a subset of international conflict featuring state or state-like actors. Second, there needs to be an initial period of active, armed conflict followed by a cessation of large-scale violent hostilities. Isolated, sporadic incidences of violence may continue to occur; however, conflict actors are either unwilling or unable to launch major military offensives. Although not necessary, this may be characterized by the establishment of a general ceasefire and/or presence of external peacekeeping troops. Third, the core issue underlying the conflict needs to remain salient post-conflict. This entails actors maintaining their claims within a dispute. And, finally, there must be an ongoing threat of violent re-escalation. In other words, post-conflict relations should be characterized by mutual hostility and an unwillingness to abandon violence as a political strategy.

These are precise criteria that expand the coverage of frozen conflicts both temporally and geographically, beyond the traditional scope of the post-Soviet separatist states that emerged in the early 1990s following the collapse of the Soviet Union. Furthermore, they hint at the dynamic nature of frozen conflicts --- far from a static condition, when a conflict freezes it may still be characterized by low levels of violence and periodic attempts at negotiations and international mediation. However, I deviate from these criteria in one crucial regard. I relax the assumption that frozen conflict is a subset of international conflict and assert that frozen conflict is a subset of conflict more broadly speaking. In regards to civil conflicts, the @smetana_between_2019 criteria restrict frozen conflicts to *de facto* independent states. This precludes the possibility that non-secessionist conflicts can exhibit the same post-conflict characteristics of unstable peace that define frozen conflicts.

Consider the case of the Second Libyan Civil War. In the aftermath of the overthrow of Muammar Gaddafi in 2011, central political authority fragmented and eventually within the midst of a renewed civil war following the contested June 2014 parliamentary elections two competing governments emerged: the internationally recognized Libyan House of Representatives (HoR) elected in 2014 based in Tubruk and in control of much of the Eastern portion of the country with the support of the Libyan National Army under Khalifa Haftar, and the Tripoli-based General National Congress (GNC) originally elected in 2012 and unilaterally self-reinstated in 2014 [@noauthor_libya_nodate-1]. A ceasefire was signed in 2015 followed by the UN brokered Libyan Political Agreement in December that established the Government of National Accord (GNA) with the intention of uniting both factions [@noauthor_libyan_2016]. The agreement was short-lived, by August the HoR had voted to reject the GNA, and cooperation rapidly deteriorated [@noauthor_libya_nodate]. Dyadic hostilities, however, remained at a low-level with only intermittent escalations while focus shifted towards the various Islamist insurgent groups also active in the conflict. Meanwhile, the core contested issue, legitimized control of the central government, remained unresolved with a strong potential for a renewal in high-intensity fighting, which occurred in 2019 with Khalifa Haftar's offensive against Tripoli and the UN-backed GNA [@hibrawi_can_2019].

I contend that the intervening years between the initial ceasefire agreement and the 2019 offensive are best characterized as a dyadic frozen conflict between non-separatist entities within a larger civil war. Although significantly shorter than the typical examples of frozen conflicts involving *de facto* independent separatist states, it satisfies the criteria of an initial period of violence followed by a cessation of large-scale hostilities, albeit with intermittent fighting, and the core issue clearly remained highly salient whilst central government authority was seriously contested.

Finally, it is worth distinguishing frozen conflict from several related concepts, specifically protracted stalemates and enduring internal rivalries. In the case of the former, stalemates during a conflict can potentially endure for extended periods of time and offer conditions similar to those of a frozen conflict. However, they do not necessitate a stop in active fighting. Instead, they may develop into wars of attrition where actors are unable to achieve an outright military advantage, but continue fighting in the hope of outlasting and "grinding down" the enemy, *i.e.* being better able to endure the prolonged costs of conflict. In the case of the latter, enduring internal rivalries are defined by @derouen_enduring_2008 as "...internal conflicts between a government and an insurgency with at least 10 years of [non-consecutive] armed conflict in which there are at least 25 deaths", an extension of the interstate concept of enduring rivalries [@goertz_enduring_1993] characterized by repeated onsets of militarized disputes. Compared to frozen conflicts, enduring internal rivalries include an explicit temporal requirement and, similar to stalemates, are not restricted to conflicts where fighting has largely stopped. The concept of frozen concepts uniquely captures the post-war condition of unstable peace that characterizes certain conflicts where there is no conclusive termination to the initial episode of fighting.

## The Emergence of Frozen Conflicts

I view the emergence of frozen conflicts as an equilibrium state of a situation where actors strictly prefer a military victory, but are unable to achieve a significant military advantage and are unwilling to acquiesce to a comprehensive peace agreement. Faced with a stalemate, belligerents will seek to minimize costs by de-escalating conflict intensity whilst maintaining a minimum degree of hostility. This may be done informally or formally --- the latter entailing ceasefires or partial peace agreements --- resulting in a cessation of large-scale active fighting following which the conflict will freeze.

To achieve this equilibrium state requires firstly a convergence in beliefs over the utility of additional fighting. Actors must believe that there is a high probability that additional fighting will simply maintain the status quo, which will decrease the incentive towards continued violence.[^convergence] From the previous definition, frozen conflict is a *post-conflict* condition that follows after the termination of the initial violent conflict. Thus, a stalemate can be considered a minimum necessary condition for the emergence of a frozen conflict if it incentivizes a conflict termination.[^stalemates]

[^stalemates]: Notably, I do not define stalemates strictly in terms of power parity since there are a number of reasons for which an actor may be unable to achieve a military advantage during the course of a conflict. This can, of course, include a balance in the material distribution of power, but also constraints arising from, for example, domestic audience costs and political interests of external, third party actors.

[^convergence]: This is akin to the "Principle of Converge" laid forth by @slantchev_principle_2003 whereby the usefulness of fighting decreases when it ceases to convey information on an opponent's capabilities and resolve. However, rather than emphasising the opportunity for bargaining as @slantchev_principle_2003, I instead link the convergence of beliefs to also the possibility for a conflict to become frozen.

Secondly, there is an implicit assumption that conflict costs are dynamic over time. Rather than conceiving of the costs of fighting as a static quantity that accumulates with each time period, I view costs as a function of the strategic choices taken by actors that influence conflict intensity. It is often assumed that one of the primary reasons for conflict termination is when the costs of fighting exceeds the expected utility of victory [@t_david_mason_how_1996].[^lit] When viewed in dynamic terms, actors have varying degrees of agency in setting costs such that a violent dispute can be maintained at a minimum level for extended periods of time. In other words, even if the expected payoff from maintaining a violent dispute is negative, actors may still avoid a settlement conditional on there being a positive future value associated with winning the dispute and no one belligerent has such a dominant military position during the conflict to effectively set the costs imposed on all other actors.

[^lit]: This assumption is often carried on in more recent literature as well. See for example the overviews provided by @reiter_exploring_2003 and @walter_bargaining_2009.

Finally, for a frozen conflict to emerge as a equilibrium state the underlying issue has to remain salient. This was previously discussed in the definition for frozen conflict, but it is worth reiterating in order to emphasize that a stalemate situation is not sufficient to guarantee that a frozen conflict will emerge. For example, in the case where a conflict terminates in an indeterminate state, but one of the belligerents unilaterally renounces violence then the underlying incompatibility as an impetus for armed conflict would no longer be pertinent. Such a scenario could hardly be described as an ongoing conflict that has frozen.

Taken together, convergent beliefs, dynamic costs, and issue salience represent minimum requirements for a conflict to become frozen, but there is still the open question, why would an actor in a stalemate situation not prefer a bargained solution terminating the conflict?[^mirror] Actors may be able to minimize costs by de-escalating the intensity; however, because the conflict is not conclusively terminated there is an ever present threat of violent re-escalation meaning that actors need to maintain some degree of military preparedness, lest the balance of power swings too far in favor of the opposing side. More formally, the expected utility of maintaining the conflict in a frozen state, with the potential to violently relitigate the issue at some future time, has to outweigh the costs associated with a frozen conflict. This means that an actor either expects to be able to gain the military advantage they were unable to achieve in the first round of fighting, or believes that they can achieve a more favourable bargain than what is currently available. To this end, I propose three different mechanisms that can plausibly incentivize actors to forgo a peaceful settlement under the condition of being unable to achieve a one-sided military victory.

[^mirror]: This mirrors the foundational question that is often the starting point for exploring the causes for the onset of war, "why fight?" If fighting is costly, then there should always exist an *ex ante* bargain dividing the contested good without incurring the cost of conflict [@fearon_rationalist_1995]. A theoretical mechanism explaining the onset of war needs to be able to account for bargaining failure. In the context of frozen conflicts, although fighting has already occurred, a complete theory needs to similarly account for why actors prefer not to seek a bargained solution with a conclusive termination instead of perpetuating the dispute.

#### Credible Commitments

That certain actors are at times unable to credibly commit to a negotiated settlement has long been identified as a potential hindrance to civil conflict termination [@walter_critical_1997; @sawyer_role_2017]. If an actor believes that an opponent will eventually renege on an agreement, they may be incentivized to fight even if they prefer to avoid war. A similar mechanism can be applied to frozen conflicts. Distrust and uncertainty can make it difficult to establish a bargaining space between actors. If neither actor is willing to continue fighting, then it may be preferable to let the conflict freeze if the costs of a frozen conflict are outweighed by the risk of the opponent reneging.

#### Transitory War Gains

A second important mechanism relates to the wartime distribution of disputed goods. During the course of a conflict, actors achieve transitory gains, such as for example territory or access to specific resources, which may not carry over into the post-conflict order under a bargained solution. If those transitory gains outweigh the benefits of a settlement, an actor who discounts the future will seek to maintain the *intrabellum* distribution and forestall any negotiations [@spaniel_stall_2020]. This is particularly applicable to frozen conflicts since by definition the belligerents are unable to agree on a bargained solution to the conflict. When the conflict freezes this also freezes the current status quo such that actors are able to continue deriving utility from their transitory war gains. Therefore, a frozen conflict may arise when there is a mismatch between the wartime distribution and the expected distribution under a post-conflict settlement.

#### Time Horizons

Lastly, actors may also prefer the utility gained from a frozen conflict over a negotiated settlement due to long time horizons. The mechanism for transitory war gains posited that actors that discount the future and value the present distribution of goods may be in favour of a frozen conflict. Although seemingly contradictory, the opposite relationship may also be present whereby an actor values the future under the expectation that they will eventually achieve a military advantage, while discounting the present costs associated with a frozen conflict. More specifically, an actor may allow a conflict to freeze to bide their time rebuilding their forces and re-strategizing in anticipation of a future violent showdown. Thus, a frozen conflict becomes a waiting device for actors unable to achieve an immediate military victory, but hoping the there will be a future favourable shift in the balance of power.[^ceasefires]

[^ceasefires]: This argument mirrors some of the work on the role of ceasefires in bargaining [@clayton_logic_2021; @sticher_bargaining_2021; @sticher_healing_2022]. Actors may accept a ceasefires according to a variety of logics, for example, they may see it as the first step towards a peace agreement, or an opportunity to gather their strength before pursuing once again a military victory.

### Distinguishing Between Civil Conflict Types

All three of these mechanisms may be present in different conflicts that have become frozen; their logics are not necessarily mutually exclusive and in some ways can be complementary. The degree to which each mechanism is relevant to any specific conflict depends in part on the conflict type. It is useful here to distinguish between civil conflicts fought over control of the state and civil conflicts fought over autonomy and/or independence claims. In the case of the former, the ultimate goal is to seize power from an incumbent regime. The prime impediment to a settlement is the risk that the post-conflict incumbent will violate the agreement, especially within authoritarian contexts where there is a fear of retribution. Thus, the preference for a frozen conflict, assuming that neither side can achieve a military advantage, may primarily stem from credible commitment problems.

In the case of the latter, civil conflicts fought over autonomy and/or independence claims, can also involve credible commitment issues as an incentive towards frozen conflict. Specifically, it is difficult for the government to credible commit to an agreement granting greater autonomy, in part because it may galvanize additional claims from other groups [@walter_information_2006; @toft_self-determination_2012]. Furthermore, as some scholars have argued, the international community tends to be fairly conservative regarding new sovereignty claims [@huddleston_foulweather_2021; @siroky_unilateral_2021]. Since rebel groups are typically outmatched by the government [@brandt_when_2008] such that it is unlikely that they will win in an absolute war and successfully disarm the regime, a frozen conflict may be the optimal outcome, especially when the opposition is able to seize and consolidate territory. If the rebel group is unable to trust the government to credibly bargain, then they will highly value the present transitory gains and seek to maintain the status quo. Meanwhile, the government may also face an incentive towards a frozen conflict in these scenarios. There is a strong international norm in favor territorial integrity.[^IR] If the government is unable to outright defeat a rebel group, especially one that has established a *de facto* independent state, then the government may be seeking a long-term plan for forcible reintegration. A frozen conflict allows the government to ease the burden of active fighting while maintaining their sovereignty claims on breakaway territory. Crucially, this latter point allows the government to frame the frozen conflict as an ongoing civil war if and when there is another violent re-escalation, as opposed to an interstate war of territorial conquest against a newly independent state.

[^IR]: I need to check the IR literature to provide a citation for this claim.

### External State Support

Where does external support by third party states to belligerents fit into the previously mentioned mechanisms and how does external support shape the utility calculation of actors either in favor or against a conflict freezing? I see external state support affecting the propensity for a frozen conflict in several different ways. Principally, support to a conflict actor can increase their military capabilities and shift the balance of power. As some scholars have argued, this is especially true in the case of directed support to rebel groups [@balch-lindsay_third-party_2008; @gent_going_2008], providing them an opportunity to launch a credible challenge to the government. This does not mean that the conflict will automatically become frozen, but it may raise the likelihood of a stalemate which is an important first step. Furthermore, external support also helps subsidize the potential cost of maintaining a frozen conflict, which may allow actors a longer time horizon such that they will reject an immediate settlement if they are optimistic about the future utility of renewed fighting.

Scholars have also argued that the involvement of additional third party actors can complicate the bargaining process for an ongoing civil conflict [@cunningham_veto_2006; @cunningham_blocking_2010]. External supporters may have conflicting outcome preferences to the actors they are supporting. As the number of actors involved in a bargaining process grows, coordinating the disparate interests becomes increasingly difficult, another disincentive against achieving a negotiated settlement.

In sum, external support should increase the likelihood of a stalemate and incentivize actors to maintain the transitory status quo in the form of a frozen conflict. My empirical expectation is that the effect should be strongest when directed support is only provided to the rebels. When support is either provided to the government, or to both sides, then it is less likely that a balance of power will emerge and I assume the situation will on average favor the government. Furthermore, I also expect this correlation to be strongest for civil conflicts concerning independence or autonomy. Given how difficult it is to secure autonomy rights from a government or achieve international sovereignty recognition for separatist rebels, as well as the general desire for governments to forcibly reintegrate territories where they maintain a sovereignty claim, there exists a stronger incentive from both sides to favour a frozen outcome. External support is especially important then to provide the capacity to rebel groups to secure territory and consolidate power.

# Empirics

To empirically assess whether external state support for conflict actors affects the probability of a civil conflict episode terminating as a frozen conflict a series of bayesian probit models were run using original data on frozen civil conflicts. In what follows, I will first present the model specification followed by a discussion of the included dependent and independent variables, including operationalizations and data sources. Finally, I will present the results and discuss the implications for understanding the emergence of frozen conflicts.

## Probit Model

I model the onset of frozen civil conflicts using a hierarchical bayesian probit model. The unit of analysis is conflict episodes --- contiguous years of active fighting according  to the UCDP definition of armed conflict as a "contested incompatibility that concerns a government and/or territory where the use of armed forced between two parties, of which at least one is the government of a state, results in at least 25 battle-related deaths within a calendar year" [@gleditsch_armed_2002; @davies_organized_2023]. The outcome, $y_i \in \{0, 1\}$ for $i = 1, 2, \ldots, N$ episodes, is a binary indicator denoting whether a given episode terminates in a frozen state modeled as a function of a matrix of treatment variables, $T$, and a set of additional control covariates, $X$, such that

\begin{align*}
y_i | \pi_i & \sim \text{Bernoulli}(\pi_i) \\
\pi_i & = \Phi\left(\alpha + X_i' \beta + T_i' \delta_{k_i} + \gamma_{c_i}\right)
\end{align*}

where $\Phi(\cdot)$ is the standard normal cumulative density function. Then, $\delta_k$ is the group specific vector of treatment effects. I allow the treatment effects to vary according to whether the underlying incompatibility for the given conflict episode concerns territory or control of the central government in order to account for potential differential effects, thus $\delta \in \mathbb{R}^{D \times K}$ for $D$ treatment variables. Finally, $\gamma_c$ represents a varying intercept for each country in order to capture unmodeled time invariant heterogeneity between countries.

In the case of the global intercept and slope coefficients for the control variables, $X$, I select fairly conservative, weakly informative priors given the restricted sample size and moderate number of predictors. Specifically, I set $\alpha \sim \text{N}(0, 5)$ and $\beta \sim \text{N}(0, 1)$ to improve sampling efficiency by placing less prior probability density on unlikely large values on the probit scale and to provide a degree of regularization.

For $\delta_k$ and $\gamma$ I use hierarchical priors that allow for varying effects while partially pooling information across groups to prevent overfitting. Regarding the former, this means $\delta_{jk} \sim \text{N}(\mu_j, \tau_j)$ for $j = 1, 2, \ldots, D$ where $\mu_j \sim \text{N}(0, 1)$ and $\tau_j \sim \text{N}(0, 2.5)$. And for the latter, $\gamma_c \sim \text{N}(0, \sigma)$ and $\sigma \sim \text{N}(0, 1)$. The particularly restrictive hyperprior on $\sigma$ shrinks the posterior $\gamma$ estimates towards zero and encodes a significant degree of skepticism towards country-specific deviations.[^country-intercepts]

[^country-intercepts]: The risk for overfitting is especially high for the country-specific intercepts since the number of observations per country is relatively small with the median country only experiencing two episodes within the time period under study.

Due to the non-linear nature of a probit regression model, in order to interpret the estimated association between the treatments and outcome, the target empirical estimands are the group-specific average marginal effects, *i.e.* the changes in the conditional expectation of the outcome variable given the various treatment states versus the baseline control averaged over the relevant subgroup sample.[^mem] Let $\theta$ denote the vector of all unknown parameters in the regression model, then the average marginal effect (AME) for the $j^{th}$ treatment state represented by vector $w_j$ for subsample $k$ can be approximated using $m = 1, 2, \ldots, M$ posterior draws from $p(\theta | y, X, T)$ such that

\begin{align*}
\text{AME}_{jk}^{(m)} = \frac{1}{N_k} \sum_{i \in S_k} P\left(y_i = 1 | T = w_j, X_i, \theta^{(m)}\right) - P\left(y_i = 1 | T = 0_{D}, X_i, \theta^{(m)}\right)
\end{align*}

[^mem]: An alternative approach would be to calculate the differences in the predicted probability for each treatment effect while holding the remaining covariates constant at their mean or mode. However, this would not necessarily be representative of the "typical case" in the population [@hanmer_behind_2013]. Although there is a minimal amount of non-stochastic missingness in the data I employ due to differing coding coverage between datasets resulting in dropped conflict episodes, I would argue that the sample averaged marginal effects are still more representative of how each treatment would affect the typical civil conflict.

## Data

### Measuring Frozen Conflicts

The outcome variable in my regression model is a binary indicator for whether a civil conflict terminates in a frozen state. Following the definition of frozen conflicts offered earlier, I construct this measure using a mix of automated and manual coding of intrastate and internationalized intrastate conflict terminations taken from the UCDP Conflict Termination Dataset [@kreutz_how_2010].[^protocol] Episodes were considered terminated if the total number of battle related fatalities dropped below 25 within a year. I subsequently labeled a termination as a frozen conflict if three criteria were satisfied. First, an episode could not result in a one-sided military victory. This would constitute a conclusive termination allowing the victor to set the terms of the post-conflict order. Second, the episode could not conclude with a comprehensive peace agreement that addresses the underlying core issue according to the UCDP Peace Agreement Dataset [@pettersson_organized_2019].[^partial] To account for the possibility of a failed peace agreement I further required that the comprehensive peace agreement had to last for at least one calendar year in order to disqualify an episode termination as a frozen conflict.[^strict_frozen] Finally, an episode could not result in actor death since this is a minimum necessary condition for a conflict to become frozen. For a state actor this meant that although there could be a turnover in the regime, the state could not cease to exist as a sovereign entity following the conclusion of the conflict episode. For non-state actors this entailed a continued existence for at least five years post-conflict without renouncing political violence and/or a re-escalation of violence into a new conflict episode within the same time period. This was done to exclude situations where rebel groups unilaterally renounce violence as a political strategy and either disband or transition into political parties or civil society organizations.

[^protocol]: For a detailed description of the codcing procedure see the protocol document in the supplemental materials.

[^partial]: Note, this does not exclude tha possibility that frozen conflicts can arise following a partial peace agreement or ceasefire. For example, the First Nagorno-Karabkh War was terminated with the 1994 Bishkek Protocol establishing a general ceasefire; however, the agreement failed to address the status of Nagorno-Karabakh leading to a frozen conflict with periodic re-escalations until the forcible reintegration of the territory by Azerbaijan following the 2023 offensive.

[^strict_frozen]: In the appendix I extend the analysis presented in this paper and run a series of robustness checks where I increase the number of years required for a compehensive peace agreement to be in place in order to better distingiush between frozen conflicts and short-term pauses in fighting.

```{r dataset}
dataset <- readRDS("./data/merged_data.rds")
n_frozen <- filter(dataset, frozen == 1) |> nrow()
n_terr_frozen <- filter(dataset, frozen == 1, incompatibility == 1) |> nrow()
```

In total, `r n_frozen` frozen conflicts were identified out of `r nrow(dataset)` civil conflict terminations, comprising approximately `r sprintf("%.0f", 100 * n_frozen / nrow(dataset))` percent of all observations for the period 1975 to 2017. Of the frozen conflicts, `r n_terr_frozen` were conflicts where the central incompatibility between conflict actors concerned territory rather than control of the central government. @fig-frozen shows both a world map of the countries that have experienced at least one frozen conflict within the coded time period as well as the incidence rate of frozen conflicts per year.

```{r}
#| fig-height: 2
#| label: fig-frozen
#| fig-cap: |
#|    On the left, a world map showing the geographic distribution of
#|    countries that have experienced at least one frozen conflict
#|    highlighted in blue. On the right, the incidence rate of new
#|    frozen conflicts per year.

frozen <- filter(dataset, frozen == 1) |>
    mutate(country = sub("Government of ", "", side_a),
           country = case_when(country == "Russia (Soviet Union)" ~ "Russia",
                               country == "United Kingdom" ~ "UK",
                               country == "Congo" ~ "Republic of Congo",
                               country == "Myanmar (Burma)" ~ "Myanmar",
                               country == "Bosnia-Herzegovina" ~ "Bosnia and Herzegovina",
                               country == "Serbia (Yugoslavia)" ~ "Serbia",
                               country == "DR Congo (Zaire)" ~ "Democratic Republic of the Congo",
                               T ~ country))

sub.df <- group_by(frozen, country) |> count(frozen)

world <- map_data("world") |>
    filter(region != "Antarctica") |>
    left_join(sub.df, by = c("region" = "country")) |>
    mutate(n = ifelse(is.na(n), 0, 1) |> as.factor())

p1 <- ggplot(world, aes(long, lat, map_id = region, fill = n, group = group)) +
    geom_map(map = world, color = "grey", size = 0.25) +
    scale_fill_manual(values = c("0" = "#CFCFC4", "1" = "#268bd2"), guide = F) +
    theme(legend.position = "none",
          axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          axis.ticks = element_blank(),
          axis.text = element_blank())

p2 <- filter(dataset, frozen == 1) |>
    ggplot(aes(year)) +
    geom_bar() +
    theme(axis.title.x = element_blank()) +
    ylab("Frozen Onsets")

p1 + p2

```

Compared to the only other extant coding of frozen conflicts, the Frozen Conflict Dataset (FCD) [@klosek_frozen_2021], the measure that I employ features a more expansive operationalization of the concept of frozen conflicts. Although the FCD covers a wider period from 1946 until 2011, only civil conflicts involving *de facto* states according to the coding by @florea_facto_2014 are included resulting in 17 cases of frozen conflict. Furthermore, the termination year for a civil conflict episode in the FCD is coded as the last year in which the number of battle related deaths is above 500. In contrast, my measure looks beyond only *de facto* states and considers civil conflicts where the central incompatibility may center over control of the government. Furthermore, I follow the UCDP and employ a smaller battle related fatalities threshold thereby capturing lower intensity conflict episodes.[^insurgencies]

[^insurgencies]: There is a risk that including low-level conflict episodes risks conflating minor, long-running insurgencies with frozen conflicts. It can be argued that the former may persist at such a low level of activity with only sporadic, isolated incidences of limited violence as to not represent a meaningfully salient issue to the incumbent state and would therefore not constitute a frozen conflict. To mitigate this risk, I also run the analysis presented in this paper on a sub-sample of conflicts with a cumulative intensity of at least 1,000 battle related fatalities since onset.

### Treatments

The primary independent variables in the model are two binary treatment variables measuring respectively whether external support was provided at least once to either the rebel side or the government side at any time during a civil conflict episode by a third party state actor, as well as the interaction in order to account for a potential differential effect when both sides receive support. Data on external state support was drawn from the UCDP External Support Dataset (ESD) [@meier_external_2023] covering the period from 1975 to 2017 and included both fungible and non-fungible support across all support categories coded by the ESD. As detailed in the model specification, I allow the effect parameters to vary according to whether the underlying incompatibility concerns control of the government or territory resulting in two sets of treatment effects.

@fig-prop shows the total counts as well as proportion of observations that terminated as frozen conflicts for each corresponding treatment category: external support to only the government, external support to only the rebels, external support to both, and no external support received by either side. In general, a higher proportion of civil conflicts centered on a territorial incompatibility terminate as a frozen conflict when support is either given only to the rebels or to both sides.

```{r}
#| fig-height: 3
#| label: fig-prop
#| fig-cap: |
#|     Total counts and proportion of observations that terminated as
#|     frozen conflict per treatment category.

conditions <- dataset |>
    mutate(treatment = case_when(ext_sup_s_rebel_bin == 1 & ext_sup_s_state_bin == 1 ~ "Both",
                                 ext_sup_s_rebel_bin == 1 & ext_sup_s_state_bin == 0 ~ "Rebel",
                                 ext_sup_s_rebel_bin == 0 & ext_sup_s_state_bin == 1 ~ "Gov.",
                                 ext_sup_s_rebel_bin == 0 & ext_sup_s_state_bin == 0 ~ "None"),
           treatment = factor(treatment, levels = c("Gov.", "Rebel", "Both", "None")),
           Incompatibility = ifelse(incompatibility == 0, "Government",
                                    "Territory"))

p1 <- ggplot(conditions, aes(treatment, fill = Incompatibility)) +
    geom_bar(position = "dodge") +
    ylab("Total Count") +
    scale_fill_solarized()

condition_props <- group_by(conditions, treatment, Incompatibility) |>
    summarise(prop = mean(frozen == 1))

p2 <- ggplot(condition_props, aes(treatment, prop, fill = Incompatibility)) +
    geom_col(position = "dodge") +
    ylim(c(0, 1)) +
    ylab("Proportion Frozen") +
    scale_fill_solarized()

p1 + p2 + plot_layout(axis_title = "collect", guides = "collect") & xlab("External Support")

```

### Control Variables

In addition to the treatment variables I also include a set of covariates that could plausibly confound the relationship between external state support and frozen conflicts. This includes variables related to  conflict dynamics, specifically: conflict episode duration, episode recurrence within a conflict, and cumulative intensity. Conflicts which last longer may affect the willingness of third party actors to offer support to belligerents. Simultaneously, long-running conflicts may also be at a greater risk of becoming frozen since the protracted nature of the conflict indicates an inability for either side to achieve a one-sided victory or be willing to reach a compromise through a peace agreement. We could expect the costs to increase such that active fighting ceases; however, the underlying conflict remains unresolved and the conflict transitions to being frozen. Thus, I include the number of years for which each episodes lasts until termination transformed with the natural logarithm function in order to reflect the diminishing impact of each one unit increase. A similar logic may also apply to episode recurrence. Recurring episodes can also indicate an intractable long-running conflict that may affect both external support and the likelihood for a conflict to become frozen. Therefore, I include a binary indicator variable for whether an episode is a recurring episode within a larger conflict as identified by the UCDP. Finally, I also include a binary indicator for whether the cumulative number of battle related deaths has exceeded 1,000 since onset. High intensity conflicts would be indicative of a greater relative strength for the opposition and may be more likely to attract outside intervention both because of the potential destabilizing effects to international/regional security as well as the greater opportunity for third party state rivals to secure a favourable post-war outcome. Furthermore, a stronger opposition may also raise the probability that the conflict becomes frozen since the incumbent government may be less likely to win an outright victory.

In complement to the partially pooled varying treatment effects, I also include a binary indicator for whether the conflict incompatibility concerns territory in order to account for a potential differing baseline probability for termination as a frozen conflict and the possibility that the conflict incompatibility affects the likelihood of external support. Further, I include variables to control for the Cold War and ongoing inter/intra-state conflicts within the host country. The former is a binary indicator for whether the termination year falls before 1989 to capture temporal dynamics related to the propensity for different types of conflict terminations. For example, international norms concerning external intervention in civil conflicts may be shaped in part by the state of the international system influencing both the decision by third party actors to intervene in a civil conflict and the types of conflict terminations that external interveners seek [@howard_how_2017]. The latter are binary indicators as well measuring whether an interstate or intrastate conflict occurred at any time during the conflict episode. Third party states involved in one conflict may be more likely to also intervene in another ongoing conflict within the same country. Furthermore, multiple ongoing conflicts could also indicate a weakened incumbent government forced to face multiple opponents, which may in turn affect the probability that a conflict freezes. Lastly, I include a binary variable for whether a conflict is left censored, *i.e.*, that the onset of the episode occurred before the coding start for the dataset, since data on external support from the ESD is only available from 1975 until 2017.[^robust_censor]

[^robust_censor]: I also run multiple robustness checks in the appendix where I alternatively drop episodes with an episode onset year before the coding start for the dataset and/or recurrent episodes within conflicts that also began prior to 1975.

## Results

@fig-ame plots the average marginal effects from estimating the hierarchical probit model predicting frozen civil conflicts given the observed data on directed external support by third party states and the set of control variables detailed in the data section.[^stan] The average marginal effects capture the change in the predicted probability for the outcome given the three treatment conditions --- external support given only to the government, external support given only to the rebel(s), and external support given to both sides --- versus the baseline condition of no external support given to any conflict actor. The plots also depict the 50% and 90% credible intervals formed from the corresponding quantiles of the posterior draws with the point estimates representing the posterior medians.

[^stan]: The model is implemented in Stan, a probabilistic programming language, and estimated using Stan's No-U-Turn sampler via the `CmdStanR` interface [@stan_development_team_stan_2024; @gabry_cmdstanr_2024]. It was run with 4 parallel chains at 2,000 iterations each with a warmup of 1,000 with convergence assessed using the standard Stan diagnostics. All pre/post-processing was done using R `r sprintf("%s.%s", R.version$major, R.version$minor)` [@r_core_team_r_2023].

```{r}
#| label: fig-ame
#| fig-height: 3
#| fig-cap: |
#|     Posterior densities for the average marginal effects for the
#|     three treatment conditions varying by incompatibility type on
#|     the predicted probability of a
#|     frozen conflict onset versus the baseline with 50% and 90%
#|     credible intervals. Point estimates are the posterior medians.

main_model <- filter(design, outcome == "frozen", treatment == "bin", episodes == "all",
                     censored == "with_censored") |> pull(id)

fit <- sprintf("posteriors/%s/fit.rds", main_model) |>  readRDS()

ame <- fit$draws("ame", format = "df") |>
    select(matches("ame\\[[124]")) |>
    pivot_longer(everything(), names_to = "variable")

ll <- extract_digits(ame$variable)
ame$variable <- sprintf("ame[%s]", sapply(ll, `[`, 1))
ame$group <- ifelse(sapply(ll, `[`, 2) == 1, "Government Incompatibility",
                    "Territorial Incompatibility")

ame2 <- filter(ame, variable == "ame[2]", group == "Territorial Incompatibility")
ame4 <- filter(ame, variable == "ame[4]", group == "Territorial Incompatibility")

ggplot(ame, aes(x = value, y = variable, fill = variable)) +
    stat_eye(alpha = 0.8, show.legend = F, .width = c(0.5, 0.9)) +
    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.2) +
    labs(x = parse(text = "bold('AME '*Delta)"),
         y = "External Support") +
    scale_y_discrete(labels = c("ame[1]" = "Gov. Only",
                                "ame[2]" = "Rebel Only",
                                "ame[4]" = "Both")) +
    coord_flip() +
    facet_grid(. ~ group) +
    scale_fill_solarized()
```

From the results, it is clear that only in instances of territorial civil conflict where external support was provided to either both sides or to only the rebel(s) is there a substantive effect on the probability of an episode becoming a frozen conflict. In both cases, the effect is positive with respective probabilities `r mean(ame4$value > 0) |> signif(2)` and `r mean(ame2$value > 0) |> signif(2)`, and on average, each condition is associated with an approximately `r mean(ame2$value) |> signif(2)` increase in the predicted probability for a frozen conflict across the sample of territorial civil conflict episodes, albeit with a high degree of posterior uncertainty, $\text{CI}_{90\%}: [`r quantile(ame4$value, 0.05) |> signif(2)`, `r quantile(ame4$value, 0.95) |> signif(2)`]$ and $\text{CI}_{90\%}: [`r quantile(ame2$value, 0.05) |> signif(2)`, `r quantile(ame2$value, 0.95) |> signif(2)`]$.

```{r}
#| label: fig-delta
#| fig-height: 3
#| fig-cap: |
#|     Posterior densities as well as 50% and 90% credible intervals
#|     for the treatment parameters varying by incompatibility type in
#|     the hierarchical probit model predicting whether a civil
#|     conflict episodes terminates as a frozen conflict. Point
#|     estimates are the posterior medians.

delta <- fit$draws("delta", format = "df") |>
    select(matches("delta")) |>
    pivot_longer(everything(), names_to = "variable")

ll <- extract_digits(delta$variable)
delta$variable <- sprintf("delta[%s]", sapply(ll, `[`, 1))
delta$group <- ifelse(sapply(ll, `[`, 2) == 1, "Government Incompatibility",
                      "Territorial Incompatibility")

delta2 <- filter(delta, variable == "delta[2]", group == "Territorial Incompatibility")

ggplot(delta, aes(x = value, y = variable, fill = variable)) +
    stat_halfeye(alpha = 0.8, show.legend = F, .width = c(0.5, 0.9)) +
    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.2) +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank()) +
    scale_y_discrete(labels = c("delta[1]" = "Ext. Sup. to Gov",
                                "delta[2]" = "Ext. Sup. to Rebel",
                                "delta[3]" = "Interaction")) +
    facet_grid(. ~ group) +
    scale_fill_solarized()
```

From the parameter plots in @fig-delta, the main driver for this substantive results is the estimate for external support given to the rebel(s) in the context of a territorial civil conflict. In this case, the probability for the treatment parameter being greater than zero is `r mean(delta2$value > 0) |> signif(2)` with $\text{CI}_{90\%}: [`r quantile(delta2$value, 0.05) |> signif(2)`, `r quantile(delta2$value, 0.95) |> signif(2)`]$. Taken together, these results indicate that the crucial predictor for whether an intra-state conflict terminates as a frozen conflict is whether the opposition received external support by a third party state during a conflict episode centered around a territorial incompatibility. Meanwhile, external support directed to the government is fairly uninformative as are all treatment effects under the condition of a government incompatibility. However, the posterior uncertainty remains high, which limits the degree to which the treatment effect can be substantively interpreted.

# Conclusion

This paper sought to investigate a potential link between external support offered by third party states and the likelihood that a civil conflict transitions into a frozen conflict. A theory was offered emphasizing the expected utility of actors
and how external support can shape the cost-benefit calculation in favour of a frozen outcome. This was then investigated empirically with a bayesian hierarchical probit model showing correlationary evidence that directed support to rebel groups does increase the probability that a civil conflict freezes, but only within the context of territorial civil conflicts.

One potential shortcoming is that I did not take into account the preferences of external supporters and motivation for offering support. Nor did I disaggregate between different types of external support. In line with other empirical work on the disaggregated effects of external support and conflict duration, there may be a differential effect depending on the type and amount of support offered to conflict belligerents.

Furthermore, the measurement offered for frozen conflicts was necessarily crude. It operationalized frozen conflict as a discrete event, "frozeness" maybe is best conceptualized as lying on a continuous dimension that allows conflicts to fluctuate  between "more" and "less" frozen. Future work should focus on refining this measure and to hopefully further expand on our understanding of both frozen conflicts and civil conflict termination.

\newpage

# References

::: {#refs}
:::

\newpage

# Appendix

## List of Frozen Conflicts

```{r}
#| label: tbl-frozen

filter(dataset, frozen == 1) |>
    arrange(side_a, year) |>
    select(Government = side_a, `Rebel Group(s)` = side_b, Year = year) |>
    mutate(Government = sub("Government of", "", Government) |> trimws()) |>
    kable(longtable = T)
```

\newpage

## Simulation Based Calibration

```{R}
#| label: fig-sbc
#| fig-cap: |
#|     Simulation based calibration results

sbc <- readRDS("./posteriors/sbc.rds") |>
    select(-.chain, -.iteration, -.draw)

# I really, really hate ggplot histograms
par(family = "TeX Gyre Pagella", mfrow = c(5, 5),
    mai = c(0.35, 0.2, 0.1, 0.2), mgp = c(1.75, 0.5, 0))
for (i in seq_along(sbc)) {
    # I also hate tibbles with every fiber of my being
    label <- colnames(sbc)[i] |> stanvar_latex()
    sbc[[i]] |> hist(main = "", xlab = TeX(label), ylab = "")
}
```

\newpage

## Robustness Checks

Different combinations of the following...

- Aggregation of the treatments: binary treatment, binary treatment in the last 5 years of the conflict episode, and proportional treatment. The latter is calculated as the proportion of conflict-years in an episode where one side received external support

- All conflict episodes and only high intensity episodes --- try to distinguish between low-level insurgencies and actual civil wars

- Normal definition for frozen conflicts and a strict definition that requires a duration of at least 2 years in order to exclude temporary one-year pauses in fighting that may be more the result of UCDP's coding decision rather than representing actual frozen conflict.

```{R}
#| label: fig-robust
#| fig-cap: Robustness checks

models <- list.dirs("posteriors", recursive = F)
stopifnot(length(models) != 0)

fits <- lapply(models, \(dir) file.path(dir, "fit.rds") |> readRDS())
names(fits) <- basename(models)

model_ames <- lapply(names(fits), function(model) {
    fits[[model]]$summary("ame", pzero, posterior::quantile2, median) |>
                    mutate(id = sub("data_", "", model)) |>
                    filter(grepl("ame\\[[124]", variable))
}) |>
    bind_rows() |>
    left_join(design, by = "id")

# Assign the correct variable names/groups
ll <- extract_digits(model_ames$variable)
ame_keys <- sprintf("ame[%s]", sapply(ll, `[`, 1))

model_ames$variable <- case_when(ame_keys == "ame[1]" ~ "Ext. Sup. Gov Only",
                                 ame_keys == "ame[2]" ~ "Ext. Sup. Rebel Only",
                                 ame_keys == "ame[4]" ~ "Ext. Sup. to Both")
model_ames$group <- ifelse(sapply(ll, `[`, 2) == 1, "Government Incompatibility",
                           "Territorial Incompatibility")

ggplot(model_ames, aes(pzero, median)) +
    geom_point(alpha = 0.2) +
    geom_errorbar(aes(ymin = q5, ymax = q95), alpha = 0.25) +
    geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.2) +
    facet_grid(variable ~ group) +
    labs(y = parse(text = "tilde(Delta)"), x = parse(text = "P(tilde(Delta) > 0)")) +
    theme(legend.position = "none")

```